---
title: "MonteCarlo test"
author: "Juan A. Dominguez J."
date: "19/9/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introducion

This document contains the info related to the session about some introductory examples for understanding the Montecarlo simulation method. This one is a tool to learn basic concepts in statistics & probability.

The Montecarlo method falls under the rubric of **Parametric bootstrap** that is quite different than the **non-parametric bootstrap**, which uses resampling of the original data to generate empirical sampling distributions


#####  Importing the library
```{r,warning =FALSE,comment='', tidy=TRUE, message=FALSE}
library(MonteCarlo)
````

#####  Generating random numbers

Generating 1000 random numbers by using one of the many useful facilities that R offer to us.

```{r message=FALSE, eval=FALSE}
runif(n=1000,min=0,max=1)
```

Ok, so now is pretty interesant to look the behaviour of these generated data. By using the `par()` function
```{r}
#If we wanted to look at 100 draws from a normal distribution with u=5 and std=2, e.g. N(5,2)
random.normal.100 <- rnorm(n=100,mean=5,sd=2)
#The par function is useful when we need to modify low-level plotting funcionts
par(mfrow=c(1,3))
plot(random.normal.100)
boxplot(random.normal.100)
hist(random.normal.100)
```

##### Look at the mean and std computed 
```{r}
mean(random.normal.100)
sd(random.normal.100)
````
*Results are closed to the theoretical ones establisteds*

-------
Now if we want to make more than just one simulation, let's use the `replicate` fucntion. In this case, we will replicate them by 5 times.
```{r}
random.normal.100.rep <- replicate(n=5,rnorm(100,5,2))
````

Several times!!
```{r eval=FALSE,}
par(mfrow=c(2,2))
#Apply() allow us to call several times one specific function (hist, mean, sd) by columns (MARGIN = 2), if this one would 1, it means Rows.
apply(X=random.normal.100.rep, MARGIN = 2, FUN= hist)
apply(X=random.normal.100.rep, MARGIN = 2, FUN= mean)
apply(X=random.normal.100.rep, MARGIN = 2, FUN= sd)
````

Summarizing them
```{r }
summary(random.normal.100.rep)
````

So, let's compute the SD among these results

````{r comment=""}
sd(apply(random.normal.100.rep,2,mean))
```
*Results show that there is not so much SD among data*

##### Let's repeat but in this case with a lower sample (25)

````{r comment=""}
nom.sim.25 <- replicate(n=5,rnorm(25,5,2))
sd(apply(nom.sim.25,2,mean))
```

#####  Let's repeat but in this case with a larger sample (1000)

````{r comment=""}
nom.sim.1000 <- replicate(n=5,rnorm(1000,5,2))
sd(apply(nom.sim.1000,2,mean))
```


At this moment, we have just learned about Montecarlo method. It was clear that when we perform simulations specially with distributions , as our n increases we will get closer to the expected value.


## Applications

##### Computing the area under the curve

Let's supose that we have a function `f(x)=1-x^2`. For computing its area, just compute the integral and define the limits.

$$
Area = \int_0^1 (1-x^2)dx\,=0.66
$$
Generating some random data
```{r}
randi<- runif(n=1000,min=0,max=1)
````

Computing the theoretical area some random data
```{r}
area.t <- 1 - randi^2
plot(randi,area.t,pch=16)
````

Generating some other random data
```{r}
randi2<- runif(n=1000,min=0,max=1)
plot(randi2,col="orange")
par(new=TRUE)
plot(randi,area.t, pch=15, col="black")
````

Computing these other random data that are belongs to the desired area

```{r}
belongs.area<-ifelse(randi2 < area.t, randi2,NA)
plot(randi2,col="orange")
par(new=TRUE)
plot(randi,belongs.area ,col="blue" )
par(new=TRUE)
plot(randi,area.t ,col="black" )

```

So, computing the number of success into the all posibles results:

```{r}
mean(randi2<area.t)
```
This result is so much closer to the theoretical value computed by the analytic solution of the integral before mentioned


##### Computing the area under the curve

# plotting the density function of a normal distribution: N(2, .25)
x <- seq(, 120, 1)

plot(x, dnorm(x, 2, 0.5), type = "l"
